---
title: "How to use the omop2survey package"
output: html_document
author: Elif Dede Yildirim
date: "2024-05-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

-   Install and import packages

```{r}
library(reticulate)
library(readr)
py_install("omop2survey")
py_install("pandas")
omop2 <- import("omop2survey")
pd <- import("pandas")
```

-   Mapping answers to numeric and text values: input data can be text, a CSV file, an Excel file, or a pandas DataFrame.

```{r}

sample_df <- read_csv("sample_survey.csv")
sample_df_copy<- sample_df
sample_df_copy<- omop2$map_answers(sample_df_copy)
head(sample_df_copy)
```

-   Create a codebook and save it as an HTML file; the codebook contains only variables in the dataset.

    Note: The codebook function can be used to save the file to the GC workspace bucket whereas codebook_html saves the file locally.

```{r}

omop2$codebook_html(sample_df_copy)

# use codebook function in the cloud environment
# omop2$codebook(sample_df_copy)
```

-   Recode missing values

```{r}
sample_recode_df <- omop2$recode(sample_df_copy)
head(sample_recode_df)
```

-   Create dummy coded variables

```{r}
sample_dummy_df <- omop2$create_dummies(sample_df_copy)
unique(sample_dummy_df$question_concept_id)
```

-   Convert data from long format to wide format using numeric values.

    Note: The pivot function can be used in the cloud environment. Use pivot_local to save files locally.

```{r}
omop2$pivot_local(sample_df_copy)

# use pivot function in the cloud environment
# omop2$pivot(sample_df_copy)
```

-   Convert data from long format to wide format using text values.

    Note: The pivot_text function can be used in the cloud environment. Use pivot_text_local to save files locally.

```{r}
omop2$pivot_text_local(sample_df_copy)

# use pivot_text function in the cloud environment
# omop2$pivot_text(sample_df_copy)
```

-   Calculate scale scores using a wide format data frame. Scale function takes following arguments; data, variables, scale_name, na =True/False, method= 'sum' / 'mean'

```{r}
#read pivot_df
pivot_df <- read_csv('workspace/pivot_n.csv')
#define variables
variables <- list('q43528662.0', 'q43528663.0', 'q43528664.0')
#define scale_name
scale_name <- 'afford_healthcare'

pivot_scale <- omop2$scale(pivot_df, variables, scale_name) # default na=False, and method='sum'
print(summary(pivot_scale$afford_healthcare))
```

```{r}
#scale scor3e using mean method
scale_name <- 'mean_afford_healthcare'
pivot_scale2 <- omop2$scale(pivot_scale, variables, scale_name, method='mean')
print(summary(pivot_scale2$mean_afford_healthcare))
```

-   Save data to the workspace

```{r}
#save the data
write_csv(pivot_scale2, 'workspace/pivot_scale.csv')
```
